A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2020-04-24 paper/paper.tex
bbl file and tex file

\documentclass{article}

@@

<<setup, include = F>>=
knitr::opts_chunk$set(echo = F, dev = "pdf", dpi = 1000, message = F, warning = F, fig.align = "center", fig.path = "figure/")
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
knit_hooks$set(document = function(x) {sub('\\usepackage[]{color}', '\\usepackage[table]{xcolor}', x, fixed = TRUE)})

library(magrittr)
library(tidyverse)
library(tidyr)
nest <- nest_legacy
unnest <- unnest_legacy
library(ggplot2)
library(keras)

mytheme <- theme_bw()
theme_set(mytheme)
@

<<results_setup, include = F>>=
if ("paper" %in% list.files()) {
  wd <- file.path(getwd(), "paper")
} else {
  wd <- getwd()
}

codedir <- file.path(wd, "code")
modeldir <- file.path(wd, "model")
imgdir <- file.path(wd, "images")

source(file.path(codedir, "Generate_Model_Images.R"))
source(file.path(codedir, "equalize.R"))
@

<<labeled-data-setup, include = F>>=
model_path <- file.path(modeldir, "TrainedModels")

if (grepl("!School", model_path)) {
  newest_model <- get_newest(dir = model_path, pattern = "weights.h5")
  newest_model$prefix = "vgg16_onehotaug_9class_256"
  newest_model$start_date = "2019-05-14_19_45_22"
} else {
  newest_model <- list(path = file.path(model_path, "20190514-155453"),
      base_file = "2019-05-14_19:45:22_vgg16_onehotaug_9class_256-weights.h5",
      prefix = "vgg16_onehotaug_9class_256", start_date = "2019-05-14_19:45:22",
      process_dir = "20190514-155453")
}

newest_data_file <- file.path(modeldir, "RProcessedImages",
                              newest_model$process_dir, "cropped_photos.Rdata")

load(newest_data_file)
source(file.path(codedir, "count_images.R"))
@

<<load-keras-model, include = F, eval = T, echo = F>>=
model_dir <- newest_model$path
load(list.files(model_dir, pattern = "-history.Rdata", full.names = T)[1])
load(file.path(newest_model$path,
               paste0(newest_model$start_date, "_",
                      newest_model$prefix, ".Rdata")))

model_wts_file <- file.path(newest_model$path, newest_model$base_file)
loaded_model <- set_weights(model_wts_file)
@

@@

%% Group authors per affiliation:
\author{Miranda Tilton and Susan Vanderplas}
% \address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}

\begin{document}
\maketitle
\begin{abstract}
Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process the quantity of data that would be required for such assessments without prohibitive amounts of human labor. In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development. The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level geometric features such as circles, stars, and quadrilaterals. These features are designed to mimic high-level features used by examiners to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks. The algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant step toward enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.
\end{abstract}
%
% \begin{keyword}
% Footwear; class characteristics; \mt{random match probability}; computer vision; neural networks
% \end{keyword}

@@

<<vgg16layers, fig.width = 7, fig.height = 5.75, fig.cap = "Visualizing convolutional filters requires generating images which maximally activate specific filters. This figure contains images that maximally activate a selection of filters from the convolutional blocks of VGG16. Filters from later blocks detect more complex features than those in earlier blocks. Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision\" trained on a general task to be applied to new domains.">>=

if(!file.exists(file.path(wd, "images", "unlabeled_filters.png"))) {
  imgs <- list.files(file.path(wd, "images", "vgg16layers"),
                     recursive = T, full.names = T)
  pngs <- lapply(imgs, function(x) png::readPNG(x)[46:1596, 32:1582, ])
  grobs <- lapply(pngs, grid::rasterGrob)
  filters <- gridExtra::arrangeGrob(grobs = grobs, nrow = 5)
  ggsave(filename = file.path(wd, "images", "unlabeled_filters.png"),
         plot = filters, device = "png", width = 6, height = 5.3)
}

if(!file.exists(file.path(wd, "images", "labeled_filters.png"))) {
  library(magick)
  b <- image_blank(width = 300, height = 1590)
  f <- image_read(file.path(wd, "images", "unlabeled_filters.png"))
  f_l <- image_append(c(b, f)) %>%
    image_annotate("Block 1:", location = "+60+139",
                   color = "black", size = 60) %>%
    image_annotate("Block 2:", location = "+60+457",
                   color = "black", size = 60) %>%
    image_annotate("Block 3:", location = "+60+775",
                   color = "black", size = 60) %>%
    image_annotate("Block 4:", location = "+60+1093",
                   color = "black", size = 60) %>%
    image_annotate("Block 5:", location = "+60+1411",
                   color = "black", size = 60) %>%
    image_scale(geometry = "600x")
  image_write(image = f_l, path = file.path(wd, "images", "labeled_filters.png"),
              format = "png")
}

knitr::include_graphics(file.path(wd, "images", "labeled_filters.png"))
@

@@

Thousands of outsole images were obtained from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}. After annotation, the minimum bounding rectangle of the region is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels. During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion. To date, \Sexpr{unique(ann_df$base_image) %>% length()} shoes have been labeled, yielding \Sexpr{nrow(dfunion)} multi-label images.

@@

<<class-characteristic-barchart, fig.width = 6, fig.height = 4, dpi = 300, out.width = "80%", fig.cap = "Distribution of classes in all labeled images. Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.", fig.scap = "Distribution of classes in all labeled images.">>=
annotated_imgs %>%
  mutate(label_type = str_replace(label_type, "multi", "multiple")) %>%
  mutate(labels = str_to_title(labels)) %>%
  mutate(labels = factor(labels, levels = c("Quad", "Line", "Text", "Circle",
                                            "Chevron", "Triangle", "Polygon",
                                            "Star", "Bowtie", "Other"))) %>%
  ggplot() +
  geom_bar(aes(x = labels, fill = label_type), color = "black") +
  scale_fill_manual("Labels", values = c("single" = "#6ba2b9",
                                         "multiple" = "#2e5597")) +
  coord_flip() +
  ylab("# Labeled Images") +
  xlab("")  +
  ggtitle("Current Class Distribution (All Labeled Images)") +
  mytheme +
  theme(legend.position = c(1, 1),
        legend.justification = c(1.03, 1.03),
        legend.background = element_rect(fill = "white"))
@

@@

The \Sexpr{nrow(dfunion)} images were split such that 60\% were used for training. Since the categories do not exist in equal proportion in the labeled data,
the training data were weighted by proportion during the training process to ensure that the model can identify both rare and common geometric shapes. Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the final model.

@@

<<training-accuracy, fig.width = 7.5, fig.height = 5, out.width = ".75\\textwidth", fig.cap = "Training and validation accuracy and loss for each epoch of the fitting process. Training and validation accuracy reach 89.5\\% around epoch 9. After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.", fig.scap = "Training and validation accuracy and loss during each epoch.", fig.pos = 'h'>>=
data.frame(history$metrics) %>%
  mutate(epoch = 1:n()) %>%
  gather(key = "measure", value = "value", -epoch) %>%
  mutate(Type = ifelse(str_detect(measure, "val"), "Validation", "Training"),
         measure = ifelse(str_detect(measure, "acc"), "Accuracy", "Loss")) %>%
  # bind_rows(tibble(epoch = NA, value =  .6, measure = "Accuracy", Type = "Validation")) %>%
  # bind_rows(tibble(epoch = NA, value =  .33, measure = "Loss", Type = "Validation")) %>%
  ggplot(aes(x = epoch, y = value, color = Type)) +
  geom_point() +
  geom_smooth(se = F) +
  facet_grid(measure~., scales = "free_y", switch = "both") +
  theme_bw() +
  scale_y_continuous("") +
  scale_x_continuous("Epoch") +
  ggtitle("CoNNOR Training Performance") + mytheme +
  theme(axis.title.y = element_blank(), legend.position = c(1, .5),
        legend.justification = c(1.03, -0.05),
        legend.background = element_rect(fill = "white"))
@

@@

<<overall-roc, fig.width = 5, fig.height = 5, out.width = ".5\\textwidth", fig.cap = "ROC curve showing overall model performance. Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.", fig.scap = "Overall model performance ROC curve.">>=
library(pROC)
pred_df <- as_tibble(preds) %>% gather(key = feature, value = value)
test_labs_df <- as_tibble(test_labs) %>% gather(key = feature, value = value)
whole_model_roc <- roc(test_labs_df$value, pred_df$value)

whole_model_roc_df <- tibble(tpr = whole_model_roc$sensitivities,
                             fpr = 1 - whole_model_roc$specificities,
                             thresholds = whole_model_roc$thresholds,
                             auc = whole_model_roc$auc[1]) %>%
  nest(tpr, fpr, thresholds, .key =  "roc_plot") %>%
  mutate(eer = purrr::map(roc_plot, eer))
ggplot() +
  geom_line(aes(x = fpr, y = tpr), data = unnest(whole_model_roc_df, roc_plot),
            size = 1.25) +
  geom_label(aes(x = 1, y = .07, label = sprintf("AUC: %0.2f", auc)), hjust = 1,
             vjust = -0.2, data = whole_model_roc_df) +
  geom_point(aes(x = fpr, y = tpr, color = "Equal Error Rate"),
             data = unnest(whole_model_roc_df, eer), size = 2) +
  scale_color_manual("", values = "black") +
  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1),
                     labels = c("0.0", "", "0.5", "", "1.0")) +
  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1),
                     labels = c("0.0", "", "0.5", "", "1.0")) +
  ggtitle("CoNNOR Test Set Performance (All Classes)") +
  coord_fixed() + mytheme +
  theme(legend.position = c(1, 0), legend.justification = c(1.01, -0.01),
        legend.title = element_blank(),
        legend.background = element_rect(fill = "white"))
@

<<class-roc, fig.width = 8, fig.height = 6, out.width = "\\textwidth", fig.cap = "Class-by-class ROC curves. AUC is area under the curve, a measure of overall model performance. Equal error rates (EER) are marked, indicating the position at which there is equal probability of a false positive or false negative error.", fig.scap = "Class-by-class ROC curves.">>=
aucs <- plot_onehot_roc(preds, test_labs, str_to_title(classes))
thresholds <- purrr::map_dbl(aucs$data$eer, ~.$thresholds)
aucs$data$thresholds <- thresholds

ggplot() +
  geom_line(aes(x = fpr, y = tpr), data = unnest(aucs$data, roc_plot), size = 1.25) +
  geom_label(aes(x = 1, y = 0, label = sprintf("AUC: %0.2f\nEER: %0.2f", auc, thresholds)), hjust = 1, vjust = -0.02, data = aucs$data) +
  geom_point(aes(x = fpr, y = tpr, color = "Equal Error\nRate (EER)"), data = unnest(aucs$data, eer), size = 2.5) +
  scale_color_manual("", values = "black") +
  facet_wrap(~class) +
  scale_x_continuous("False Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
  scale_y_continuous("True Positive Rate", breaks = c(0, .25, .5, .75, 1), labels = c("0.0", "", "0.5", "", "1.0")) +
  ggtitle("CoNNOR Test Set Performance") +
  facet_wrap(~class, nrow = 2) +
  coord_fixed() +
  theme(legend.position = c(1, 0), legend.justification = c(1, 0))
@

The full model has an AUC of \Sexpr{round(whole_model_roc_df$auc,2)}, and the AUC for individual classes ranges from \Sexpr{round(min(aucs$data$auc),2)} (for \Sexpr{sprintf("%s",classes[which.min(aucs$data$auc)])}) to \Sexpr{round(max(aucs$data$auc),2)}  (for \Sexpr{sprintf("%s",classes[which.max(aucs$data$auc)])}).
While the class performances do vary slightly, each ROC curve indicates that the model performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. We will use a class specific EER to set an individual threshold for each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.

@@

<<eers, include = F>>=
eers <- purrr::map_dbl(aucs$data$eer, function(x){x$thresholds}) %>%
  setNames(classes)
@

<<exemplars, fig.cap = "Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories. Predicted probabilities greater than the equal error rate threshold are outlined in blue. Although there are some misclassifications, predictions are generally consistent with the human-assigned labels of these images.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
single_exemplars <- list.files(file.path(wd, "images", "consistency",
                                         "single-exemplar"), full.names = T)
multi_exemplars <- list.files(file.path(wd, "images", "consistency",
                                        "multi-exemplar"), full.names = T)
pred_prob_plot(rev(c(single_exemplars, multi_exemplars)),
               loaded_model, sort = F, eer = eers)
@

While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency. \autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text. The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text. The remaining images contain only one of either text or circles, as well as other possible shape categories. In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final 3 images contain pure circles that do not appear to belong to text. The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having difficulty distinguishing the two categories. While the circle predictions are often above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular. This confusion is complicated by the frequency of images which contain both circles and text: text is present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("circle", dfunion$name)), 2)}\% of our labeled images containing circles, and circles are present in \Sexpr{100 * round(sum(as.logical(grepl("circle", dfunion$name) & grepl("text", dfunion$name))) / sum(grepl("text", dfunion$name)), 2)}\% of our images containing text. In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape. More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.

<<circle-text-consistency, fig.cap = "Model predictions of images containing circles and/or text. The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G\". Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast text within a hexagon and the false positive prediction of concentric circles.", fig.pos = "p!", out.width = "100%", cache = T, fig.width = 11.5/1.25, fig.height = 16/1.25>>=
circle_text <- list.files(file.path(wd, "images", "consistency", "circle-text"), full.names = T)
pred_prob_plot(rev(circle_text), loaded_model, sort = F, eer = eers)
@

@@

<<confusion-matrix, fig.width = 9, fig.height = 8, out.width = "\\textwidth", fig.cap = "Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors. Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives. Quadrilaterals are generally over-predicted across true labels; polygons and triangles are under-predicted. Circles and text are also often confused.", fig.scap = "Confusion matrix, with correct and incorrect model classifications.", dpi = 600>>=
get_confusion_matrix(predictions = preds, classes = classes,
                     test_labels = test_labs, threshold = thresholds) %>%
  set_names(str_to_title(classes)) %>%
  ggcorrplot(., hc.order = F, outline.col = "white", lab = T) +
  theme_bw() +
  scale_fill_gradient("Classification\nRate", low = "white",
                      high = "cornflowerblue", limits = c(0, 1)) +
  scale_x_discrete("Image Label", labels = str_to_title(classes)) + scale_y_discrete("Prediction") +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14, angle = 90, vjust = 1),
        axis.text.y = element_text(angle = 90, hjust = .5)) +
  ggtitle("CoNNOR Multi-Class Confusion Matrix: Test Set Performance") +
  theme(panel.grid.major = element_line(color = "grey50"),
        panel.grid.minor = element_line(color = "grey60")) +
  theme(plot.margin = grid::unit(c(1,0,0,0), "mm"),
        plot.background = element_rect(fill = "transparent", color = NA),
        plot.subtitle = element_blank(), plot.caption = element_blank(),
        panel.spacing = unit(c(0, 0, 0, 0), "mm"))
@

@@

<<low-contrast-consistency, fig.cap = "Model predictions for images with low contrast. While these images contain a variety of shapes (as shown in the color-corrected column on the right), the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.", fig.pos = "p!", out.width = "100%", fig.width = 10/1.2, fig.height = 9/1.2, cache = T>>=
lowcontrast <- list.files(file.path(wd, "images", "consistency", "low-contrast2"), full.names = T)

pred_prob_plot(lowcontrast, loaded_model, sort = F, eer = eers,
               img_path2 = lowcontrast) +
  theme(legend.position = "bottom")
@

@@

<<brand-plots, out.width = "100%", fig.width = 7, fig.height = 5, fig.cap = "Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence. Brand text is present when the brand makes up more than five percent of labeled shoes for the shape category. Some shapes, like bowtie and polygon, are relatively dominated by a small number of brands while other shapes, such as text and quad, are comprised of a large number of brands. Typically, model predictions are best for shape categories without much variability, which, in general, coincides with shapes dominated by a small number of brands.">>=
img_names <- dfunion$filename %>% basename() %>% as.vector

labs <- sapply(classes, grepl, x = str_extract(img_names, "^.*?-\\d?-"))
str <- gsub("^.*?-\\d{0,2}-", "", img_names)

# Extract first, first two, and first three words of string
brand_1 <- str_extract(str, "^.*?-")
brand_2 <- str_extract(str, "^.*?-.*?-")
brand_3 <- str_extract(str, "^.*?-.*?-.*?-")

# These are the beginning of brands that need an extra word extracted
multi_brands <- c("the", "the-north", "polo", "polo-ralph", "5", "5-11",
                  "1", "to", "dr", "la", "new", "old", "under",
                  "steve", "bernie", "cole", "tory",
                  "harley", "kristin", "eric",
                  "spring", "chinese", "dirty")

brand <- ifelse(brand_1 %in% paste0(multi_brands, "-"),
                ifelse(brand_2 %in% paste0(multi_brands, "-"),
                       brand_3, brand_2),
                brand_1) %>%
  gsub(. , pattern = "-", replacement = " ") %>%
  trimws(., "right"); rm(brand_1, brand_2, brand_3, multi_brands)

df_images <- cbind(tibble(str, brand), labs)

df_shoes <- df_images %>%
  group_by(str, brand) %>%
  summarize_at(vars(bowtie:triangle), any)

df_shoes_long <- df_shoes %>%
  tidyr::gather(key = feature, value = value, -str, -brand) %>%
  ungroup() %>%
  tidyr::nest_legacy(c(str, value)) %>%
  mutate(n_true = purrr::map_int(data, ~sum(.$value)),
         n_shoes_brand = purrr::map_int(data, ~length(.$value)))

brand_prop_by_feature <- df_shoes_long %>%
  group_by(feature) %>%
  mutate(prop = n_true/sum(n_true)) %>%
  arrange(feature, desc(prop)) %>%
  mutate(brand_cat = ifelse(row_number() <= 10, brand, "other")) %>%
  group_by(feature) %>%
  mutate(rank = pmin(row_number(), 11))

brand_prop_by_feature_10 <- brand_prop_by_feature %>%
  group_by(feature, brand_cat, rank) %>%
  summarize(prop = sum(prop)) %>%
  ungroup() %>%
  group_by(feature) %>%
  arrange(feature, rank) %>%
  mutate(cprop = cumsum(prop),
         cprop2 = lag(cprop, 1, default = 0),
         mid = (cprop + cprop2)/2,
         rank = ifelse(rank == 11, NA, rank))

ggplot(data = brand_prop_by_feature_10) +
  geom_bar(aes(x = feature, y = prop, fill = factor(rank)),
           position = "stack", stat = "identity", color = "black") +
  geom_text(aes(x = feature, y = 1-mid, label = brand_cat),
            data = filter(brand_prop_by_feature_10, prop > .05), size = 2.5) +
  scale_fill_brewer(palette = "Paired", guide = F, na.value = "grey") +
  scale_y_continuous("Proportion") + scale_x_discrete("") +
  ggtitle("Proportion of Labeled Features by Brand")
@

@@

<<generate-heatmaps, echo = F, eval = F>>=
# My computer has troubles with the transparency of the overlay...
# This code is functional, but highest quality will come from generating these
# images on Bigfoot and loading them statically.

>>>>

\documentclass{article}\usepackage[]{graphicx}\usepackage[table]{xcolor}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

@@

%% Group authors per affiliation:
\author{Miranda Tilton and Susan Vanderplas}
% \address{Center for Statistics and Applications in Forensic Evidence, Iowa State University. 195 Durham Center, 613 Morill Rd., Ames, IA 50011}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\maketitle
\begin{abstract}
Quantitative assessment of class characteristic frequency is a difficult problem, in part because there is currently no way to collect and efficiently process the quantity of data that would be required for such assessments without prohibitive amounts of human labor. In this paper, we discuss a computational method which complements a passive, unmanned scanning device that is currently under development. The automatic class characteristic recognition algorithm described in this paper has been developed with photographs of outsole images, and identifies high-level geometric features such as circles, stars, and quadrilaterals. These features are designed to mimic high-level features used by examiners to facilitate automatic processing of features without resorting to black-box feature sets commonly used in neural networks. The algorithm performs well on images with unambiguous shapes and moderate color contrast, though some systematic errors arise from the physical similarities of geometric shapes. Additional improvements may be realized by modifying the labeling scheme to reduce feature overlap and by preprocessing the images to improve contrast. This new approach represents a significant step toward enabling the collection and application of population-level class characteristic data to support quantitative assessments of the value of class-characteristic based similarity assessments.
\end{abstract}
%
% \begin{keyword}
% Footwear; class characteristics; \mt{random match probability}; computer vision; neural networks
% \end{keyword}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\maxwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/labeled_filters}

}

\caption[Visualizing convolutional filters requires generating images which maximally activate specific filters]{Visualizing convolutional filters requires generating images which maximally activate specific filters. This figure contains images that maximally activate a selection of filters from the convolutional blocks of VGG16. Filters from later blocks detect more complex features than those in earlier blocks. Much like human vision, a hierarchical structure of feature detection with increasing complexity allows CNN ``vision" trained on a general task to be applied to new domains.}\label{fig:vgg16layers}
\end{figure}

\end{knitrout}

@@

Thousands of outsole images were obtained from online shoe retail sites and annotated using LabelMe, a tool for image annotation in computer vision problems \citep{labelme}. After annotation, the minimum bounding rectangle of the region is identified and the image is cropped to that area; subsequently, the cropped image is scaled to 256 x 256 pixels. During this process, aspect ratio is not preserved, though efforts are made to label regions which are relatively square to minimize the effect of this distortion. To date, 4571 shoes have been labeled, yielding 28081 multi-label images.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{figure/class-characteristic-barchart-1}

}

\caption[Distribution of classes in all labeled images.]{Distribution of classes in all labeled images. Quadrilaterals, lines, circles, text, and chevrons are relatively common; stars, polygons, and bowties are relatively rare.}\label{fig:class-characteristic-barchart}
\end{figure}

\end{knitrout}

@@

The 28081 images were split such that 60\% were used for training. Since the categories do not exist in equal proportion in the labeled data,
the training data were weighted by proportion during the training process to ensure that the model can identify both rare and common geometric shapes. Of the remaining 40\% of data, half were used for validation, to monitor the training process, and the remaining data were for testing the performance of the final model.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[h]

{\centering \includegraphics[width=.75\textwidth]{figure/training-accuracy-1}

}

\caption[Training and validation accuracy and loss during each epoch.]{Training and validation accuracy and loss for each epoch of the fitting process. Training and validation accuracy reach 89.5\% around epoch 9. After that point, validation loss remains the same and training loss decreases slightly, while validation accuracy increases more slowly than training accuracy.}\label{fig:training-accuracy}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=.5\textwidth]{figure/overall-roc-1}

}

\caption[Overall model performance ROC curve.]{ROC curve showing overall model performance. Model accuracy is generally high, as the ROC curve is significantly above the identity line $y=x$.}\label{fig:overall-roc}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/class-roc-1}

}

\caption[Class-by-class ROC curves.]{Class-by-class ROC curves. AUC is area under the curve, a measure of overall model performance. Equal error rates (EER) are marked, indicating the position at which there is equal probability of a false positive or false negative error.}\label{fig:class-roc}
\end{figure}

\end{knitrout}

The full model has an AUC of 0.88, and the AUC for individual classes ranges from 0.81 (for line) to 0.93  (for bowtie).
While the class performances do vary slightly, each ROC curve indicates that the model performs significantly better than random chance. The points in \autoref{fig:class-roc} represent the equal error rate (EER), which is the threshold where there is equal probability of a false positive or false negative error. We will use a class specific EER to set an individual threshold for each class, so that any model prediction with probability above the EER for that shape is considered a positive prediction.

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[p!]

{\centering \includegraphics[width=1\linewidth]{figure/exemplars-1}

}

\caption[Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories]{Model predictions of images containing (from the top) a single shape, no shape, or multiple shapes across all categories. Predicted probabilities greater than the equal error rate threshold are outlined in blue. Although there are some misclassifications, predictions are generally consistent with the human-assigned labels of these images.}\label{fig:exemplars}
\end{figure}

\end{knitrout}

While figure \ref{fig:exemplars} shows representative examples of classifications across multiple shape categories, it is also useful to examine many cases within a given shape to get a sense of model consistency. \autoref{fig:circle-text-consistency} shows a number of images containing circles and/or text. The top four rows show text contained within a separate, distinct circle shape, and the next five rows show circles represented as the letter ``o" in text. The remaining images contain only one of either text or circles, as well as other possible shape categories. In particular, the first two of these images contain text that use the capital letter ``G", the next two images contain no circular shapes, and the final 3 images contain pure circles that do not appear to belong to text. The model is able to identify the circle in all cases where a circle is truly present, regardless of whether the circle is part of the text or it is distinct; however, the circle prediction is also strong in three of the four images that do not contain a circle, which indicates that the model is having difficulty distinguishing the two categories. While the circle predictions are often above the EER for letters such as G and S, these predictions are generally much lower than those where distinct circles or Os are seen, indicating that there are some indications that features are not as strong in the quasi-circle shapes, but that the model is not able to completely separate Os from C, S, and G shapes which are also very circular. This confusion is complicated by the frequency of images which contain both circles and text: text is present in 39\% of our labeled images containing circles, and circles are present in 31\% of our images containing text. In addition, the distinction between a circle and an ``o" comes from the context around the shape, which is much more difficult for the model to parse than simply detecting a shape. More importantly, this behavior does not suggest poor model performance as much as a weakness of the classification scheme, which will be discussed in a later section.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[p!]

{\centering \includegraphics[width=1\linewidth]{figure/circle-text-consistency-1}

}

\caption[Model predictions of images containing circles and/or text]{Model predictions of images containing circles and/or text. The model correctly identifies circles in all images where circles are present, but also predicts moderate probabilities of circles in many images where text is present with curved letters, such as ``G". Text is correctly identified in most images, with the exception of the false negative prediction of the low-contrast text within a hexagon and the false positive prediction of concentric circles.}\label{fig:circle-text-consistency}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figure/confusion-matrix-1}

}

\caption[Confusion matrix, with correct and incorrect model classifications.]{Confusion matrix, showing on the diagonal the correct classification rate and on the off-diagonal, classification errors. Note that in multi-label images, correct off-diagonal labels have been excluded from the calculation of false positives. Quadrilaterals are generally over-predicted across true labels; polygons and triangles are under-predicted. Circles and text are also often confused.}\label{fig:confusion-matrix}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[p!]

{\centering \includegraphics[width=1\linewidth]{figure/low-contrast-consistency-1}

}

\caption[Model predictions for images with low contrast]{Model predictions for images with low contrast. While these images contain a variety of shapes (as shown in the color-corrected column on the right), the diffuse probabilities assigned by the model across all categories indicate that features are obscured by low contrast, which propogates through model training and prediction.}\label{fig:low-contrast-consistency}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figure/brand-plots-1}

}

\caption[Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence]{Proportions of top 10 shoe brands in our labeled dataset within each shape category, sorted by prevalence. Brand text is present when the brand makes up more than five percent of labeled shoes for the shape category. Some shapes, like bowtie and polygon, are relatively dominated by a small number of brands while other shapes, such as text and quad, are comprised of a large number of brands. Typically, model predictions are best for shape categories without much variability, which, in general, coincides with shapes dominated by a small number of brands.}\label{fig:brand-plots}
\end{figure}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!p]

{\centering \includegraphics[width=.95\textwidth]{/home/susan/Projects/Work/2019-CoNNORFSI/paper/images/heatmaps/processed/heatmap-a7-Ugg-grey}

}
